---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
library(cape)
```

[![Travis build status](https://travis-ci.com/stephaneguerrier/cape.svg?branch=master)](https://travis-ci.com/stephaneguerrier/cape)
[![Licence](https://img.shields.io/badge/licence-AGPL--3.0-blue.svg)](https://opensource.org/licenses/AGPL-3.0)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.0-6666ff.svg)](https://cran.r-project.org/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/SMAC-Group/simts)

# `cape` Overview 

This R package provides an implementation of the ContionAl Prevalence Estimation (or `cape`) apporach proposed in *Accurate Prevalence Estimation for Emerging or Rare Infectious Diseases* by Stéphane Guerrier, Christoph Kuzmics and Maria-Pia Victoria-Feser (submitted manuscript available upon request).

# Package installation

The `cape` pacakge can be installed from GitHub as follows:

```{r, eval=FALSE}
# Install devtools
install.packages("devtools")

# Install the package from GitHub
devtools::install_github("stephaneguerrier/CPreval")
```

# License

The license this source code is released under is the GNU AFFERO GENERAL PUBLIC LICENSE (AGPL) v3.0. Please see the LICENSE file for full text. Otherwise, please consult [TLDR Legal](https://tldrlegal.com/license/gnu-affero-general-public-license-v3-(agpl-3.0)) or [GNU](https://www.gnu.org/licenses/agpl-3.0.en.html) which will provide a synopsis of the restrictions placed upon the code.
