---
title: "Results Reproducibility"
bibliography: bibliography.bib
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Results Reproducibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Case study: Application to Austrian COVID-19 survey

We use the methodology developed in this paper for the case of the COVID-19 prevalence estimation using the results of a survey done in November 2020 by @StatAU20. We also compare the different approaches, in order to illustrate, in practice, the impact of choosing one method rather than another one. In November 2020, a survey sample of $n=2287$ was collected to test for COVID-19 using PCR-tests. Seventy-two participants were tested positive, and among these ones, thirty-five ($R_1=35$) had declared to have been tested positive with the official procedure, during the same month. In November, there were $93914$  declared cases among the official (approximately) $7166167$ inhabitants in Austria (above 16 years old), so that $\pi_0 \approx 1.3105\%$. The sensitivity ($1-\alpha$) and the specificity ($1-\beta$) are not known with precision, so that we present estimates of the prevalence without misclassification error as well as for values for the FP and FN rates, that are plausible given the data and according to the sensitivity and specificity reported in \cite{KoWeGr:20} or \cite{surkova2020false}.

The results presented in Table 2 of @guerrier2020prevalence can be reproduced as follows:

```{r}
# Load pempi
library(pempi)

# Austrian data (November 2020)
pi0 = 93914/7166167

# Load data
data("covid19_austria")

# Random sampling
n = nrow(covid19_austria)
R1 = sum(covid19_austria$Y == 1 & covid19_austria$Z == 1)
R2 = sum(covid19_austria$Y == 0 & covid19_austria$Z == 1)
R3 = sum(covid19_austria$Y == 1 & covid19_austria$Z == 0)
R4 = sum(covid19_austria$Y == 0 & covid19_austria$Z == 0)

# Weighted sampling
R1w = sum(covid19_austria$weights[covid19_austria$Y == 1 & covid19_austria$Z == 1])
R2w = sum(covid19_austria$weights[covid19_austria$Y == 0 & covid19_austria$Z == 1])
R3w = sum(covid19_austria$weights[covid19_austria$Y == 1 & covid19_austria$Z == 0])
R4w = sum(covid19_austria$weights[covid19_austria$Y == 0 & covid19_austria$Z == 0])

# Print table
data_mat = matrix(c(R1w, R2w, R3w, R4w, R1, R2, R3, R4), 2, 4, byrow = TRUE)
rownames(data_mat) = c("Weighted sampling", "Unweighted sampling")
colnames(data_mat) = c("R1 (R11)", "R2 (R10)", "R3 (R01)", "R4 (R00)")
knitr::kable(round(data_mat, 4))
```

The data can be summarized as follows: 

- $n$ = `r n`
- $\pi_0$ = `r round(pi0*100,4)`%
- $R_{11}$ = `r R1` (which is denoted as `R1` in the package).
- $\bar{R}_{11}$ = `r round(R1w,4)` (which is denoted as `R1` or `R1w` in the package).
- $R_{10}$ = `r R2` (which is denoted as `R2` in the package).
- $\bar{R}_{10}$ = `r round(R2w,4)` (which is denoted as `R2` or `R2w` in the package).
- $R_{01}$ = `r R3` (which is denoted as `R3` in the package).
- $\bar{R}_{01}$ = `r round(R3w,4)` (which is denoted as `R3` or `R3w` in the package).
- $R_{00}$ = `r R4` (which is denoted as `R4` in the package).
- $\bar{R}_{00}$ = `r round(R4w,4)` (which is denoted as `R4` or `R4w` in the package).

We can check that $R_{11} + R_{10} + R_{01} + R_{00} = n$

```{r}
R1 + R2 + R3 + R4
```

and $\bar{R}_{11} + \bar{R}_{10} + \bar{R}_{01} + \bar{R}_{00} = n$

```{r}
R1w + R2w + R3w + R4w
```

For the analysis we consider the possibility of measurement error, in which we use the following values:

```{r}
# Measurement error
alpha0 = 0
alpha = 1/100
beta = 10/100
```

# Survey MLE

## Survey MLE with random sampling

The Survey MLE (with and without measurement error) can be compute as follows and correspond to the number report in Table 3 of @guerrier2020prevalence: 

```{r}
# Survey MLE without measurement error
(smle_no_meas_error_random = survey_mle(R = R1 + R3, n = n))

# Survey MLE with measurement error (as defined above)
(smle_with_meas_error_random = survey_mle(R = R1 + R3, n = n, 
                              alpha = alpha, beta = beta))
```

## Survey MLE with stratified sampling

In the case of a stratified sampling, the Survey MLE (with and without measurement error) can be compute as follows: 

```{r}
# Survey (weighted) MLE without measurement error
(smle_no_meas_error_strat = survey_mle(R = R1w + R3w, n = n, 
                            V = mean(covid19_austria$weights^2)))

# Survey MLE with measurement error (as defined above)
(smle_with_meas_error_strat = survey_mle(R = R1w + R3w, n = n,
                             alpha = alpha, beta = beta, 
                             V = mean(covid19_austria$weights^2)))
```

# Moment-based estimator 

## Moment-based estimator with random sampling

In the case of a random sampling, the moment-based estimator or MME (with and without measurement error) can be compute as follows: 

```{r}
# MME without measurement error
(mme_no_meas_error_random = moment_estimator(R3 = R3, n = n, 
                            pi0 = pi0))

# MME with measurement error (as defined above)
(mme_with_meas_error_random = moment_estimator(R3 = R3, n = n,
                              pi0 = pi0, alpha = alpha, 
                              beta = beta, alpha0 = alpha0))
```

## Moment-based estimator with stratified sampling

In the case of a stratified sampling, the MME (with and without measurement error) can be compute as follows and correspond to the number report in Table 3 of @guerrier2020prevalence:

```{r}
# MME without measurement error
(mme_no_meas_error_strat = moment_estimator(R3 = R3w, n = n,
                          pi0 = pi0,
                          V = mean(covid19_austria$weights^2)))

# MME with measurement error (as defined above)
(mme_with_meas_error_strat = moment_estimator(R3 = R3w, n = n,
                             pi0 = pi0, alpha = alpha, beta = beta,
                             alpha0 = alpha0, 
                             V = mean(covid19_austria$weights^2)))
```

# Conditional MLE

## Conditional MLE with random sampling

In the case of a random sampling, the conditional MLE or CMLE (with and without measurement error) can be compute as follows: 

```{r}
# CMLE without measurement error
(cmle_no_meas_error_random = conditional_mle(R1 = R1, R2 = R2, 
                            R3 = R3, R4 = R4, pi0 = pi0))

# CMLE with measurement error (as defined above)
(cmle_with_meas_error_random = conditional_mle(R1 = R1, R2 = R2, 
                              R3 = R3, R4 = R4, pi0 = pi0, 
                              alpha = alpha, beta = beta, 
                              alpha0 = alpha0))
```

## Conditional MLE with stratified sampling

In the case of a stratified sampling, the CMLE (with and without measurement error) can be compute as follows: 

```{r}
# CMLE without measurement error
(cmle_no_meas_error_strat = conditional_mle(R1 = R1w, R2 = R2w, 
                            R3 = R3w, R4 = R4w, pi0 = pi0, 
                            V = mean(covid19_austria$weights^2)))

# CMLE with measurement error (as defined above)
(cmle_with_meas_error_strat = conditional_mle(R1 = R1w, R2 = R2w, 
                              R3 = R3w, R4 = R4w, n = n, pi0 = pi0,
                              alpha = alpha, beta = beta, 
                              alpha0 = alpha0, 
                              V = mean(covid19_austria$weights^2)))
```

# Marginal MLE

## Marginal MLE with random sampling

In the case of a random sampling, the marginal MLE or MMLE (with and without measurement error) can be compute as follows: 

```{r}
# MMLE without measurement error
(mmle_no_meas_error_random = marginal_mle(R1 = R1, R3 = R3, n = n, pi0 = pi0))

# MMLE with measurement error (as defined above)
(mmle_with_meas_error_random  = marginal_mle(R1 = R1, R3 = R3, n = n, pi0 = pi0, alpha = alpha, beta = beta, alpha0 = alpha0))
```

## Marginal MLE with stratified sampling

The MMLE is currently not implemented in the case of a stratified sampling.

# Replicating Table 3

Table 3 can be replicated as follows:

```{r}
table2 = matrix(NA, 6, 6)
rownames(table2) = c("SMLE (stratified)", "MME (stratified)", "Estimated beta0 (stratified)", "SMLE (random)", "MME (random)", "Estimated beta0 (random)")
colnames(table2) = c("Estimates (no meas. err.)", "95% CI (low)", "95% CI (high)", "Estimates (with meas. err.)", "95% CI (low)", "95% CI (high)")

table2[1, ] = 100*c(smle_no_meas_error_strat$estimate,
                smle_no_meas_error_strat$ci_asym,
                smle_with_meas_error_strat$estimate,
                smle_with_meas_error_strat$ci_asym) 

table2[2, ] = 100*c(mme_no_meas_error_strat$estimate,
                mme_no_meas_error_strat$ci_asym,
                mme_with_meas_error_strat$estimate,
                mme_with_meas_error_strat$ci_asym) 

table2[3, ] = 100*c(mme_no_meas_error_strat$beta0,
                mme_no_meas_error_strat$ci_beta0,
                mme_with_meas_error_strat$beta0,
                mme_with_meas_error_strat$ci_beta0)

table2[4, ] = 100*c(smle_no_meas_error_random$estimate,
                smle_no_meas_error_random$ci_asym,
                smle_with_meas_error_random$estimate,
                smle_with_meas_error_random$ci_asym)

table2[5, ] = 100*c(mme_no_meas_error_random$estimate,
                mme_no_meas_error_random$ci_asym,
                mme_with_meas_error_random$estimate,
                mme_with_meas_error_random$ci_asym) 
table2[6, ] = 100*c(mme_no_meas_error_random$beta0,
                mme_no_meas_error_random$ci_beta0,
                mme_with_meas_error_random$beta0,
                mme_with_meas_error_random$ci_beta0)

knitr::kable(round(table2, 3))
```

# Comparing prevalence

This figure can be obtained by running the file `pempi/figures/case_study.Rnw`. A similar base R version can be obtained as follows:

```{r, fig.align='center', fig.height=5, fig.width=6}
pi0 = 93914/7166167
cols = c("#F8766DFF", "#00BFC4FF")
delta = 0.1
cex_pt = 1.5
lwd_ci = 2
pch_mme = 16
pch_smle = 15

plot(NA, xlim = c(0.75, 4.25), ylim = c(1, 4), axes = FALSE, ann = FALSE)
grid()
box()
col_text = "grey60"
cex_text = 0.85
cex_text2 = 0.65
abline(v = c(1.5, 2.5, 3.5), col = col_text)

axis(2)

mtext("Stratified sampling", side = 3, line = 1.75, cex = cex_text, at = 1, col = col_text)
mtext("no measurment error", side = 3, line = 0.75, cex = cex_text2, at = 1, col = col_text)

mtext("Random sampling", side = 3, line = 1.75, cex = cex_text, at = 2, col = col_text)
mtext("no measurment error", side = 3, line = 0.75, cex = cex_text2, at = 2, col = col_text)

mtext("Stratified sampling", side = 3, line = 1.75, cex = cex_text, at = 3, col = col_text)
mtext("with measurment error", side = 3, line = 0.75, cex = cex_text2, at = 3, col = col_text)

mtext("Random sampling", side = 3, line = 1.75, cex = cex_text, at = 4, col = col_text)
mtext("with measurment error", side = 3, line = 0.75, cex = cex_text2, at = 4, col = col_text)

mtext("Prevalence (%)", side = 2, line = 3, cex = 1.15)
abline(h = 100*pi0, lwd = 2, lty = 2)

text(1, 1.18, expression(pi[0]), cex = 1.15)

legend("topright", c("MME", "95% CI",
                    "Survey MLE", "95% CI"),
       bty = "n", col = c(cols[1], cols[1], cols[2], cols[2]),
       lwd = c(NA, lwd_ci, NA, lwd_ci), pch = c(pch_mme, NA, pch_smle, NA),
       pt.cex = 1.5, cex = 0.7)

# 1) Stratified sampling, without ME
points(1 - delta, 100*mme_no_meas_error_strat$estimate, col = cols[1], pch = pch_mme, cex = cex_pt)
lines(c(1, 1) - delta, 100*mme_no_meas_error_strat$ci_asym, col = cols[1], lwd = lwd_ci)

points(1 + delta, 100*smle_no_meas_error_strat$estimate, col = cols[2], pch = pch_smle, cex = cex_pt)
lines(c(1, 1) + delta, 100*smle_no_meas_error_strat$ci_asym, col = cols[2], lwd = lwd_ci)

# 2) Random sampling, without ME
points(2 - delta, 100*mme_no_meas_error_random$estimate, col = cols[1], pch = pch_mme, cex = cex_pt)
lines(c(2, 2) - delta, 100*mme_no_meas_error_random$ci_asym, col = cols[1], lwd = lwd_ci)

points(2 + delta, 100*smle_no_meas_error_random$estimate, col = cols[2], pch = pch_smle, cex = cex_pt)
lines(c(2, 2) + delta, 100*smle_no_meas_error_random$ci_asym, col = cols[2], lwd = lwd_ci)

# 3) Stratified sampling, with ME
points(3 - delta, 100*mme_with_meas_error_strat$estimate, col = cols[1], pch = pch_mme, cex = cex_pt)
lines(c(3, 3) - delta, 100*mme_with_meas_error_strat$ci_asym, col = cols[1], lwd = lwd_ci)

points(3 + delta, 100*smle_with_meas_error_strat$estimate, col = cols[2], pch = pch_smle, cex = cex_pt)
lines(c(3, 3) + delta, 100*smle_with_meas_error_strat$ci_asym, col = cols[2], lwd = lwd_ci)

# 4) Random sampling, with ME
points(4 - delta, 100*mme_with_meas_error_random$estimate, col = cols[1], pch = pch_mme, cex = cex_pt)
lines(c(4, 4) - delta, 100*mme_with_meas_error_random$ci_asym, col = cols[1], lwd = lwd_ci)

points(4 + delta, 100*smle_with_meas_error_random$estimate, col = cols[2], pch = pch_smle, cex = cex_pt)
lines(c(4, 4) + delta, 100*smle_with_meas_error_random$ci_asym, col = cols[2], lwd = lwd_ci)
```

# Replicating Figure 1

Figure 1 can be obtained by running the file `pempi/figures/case_study.Rnw`. A similar base R version can be obtained as follows:

```{r, fig.align='center', fig.height=5, fig.width=6}
# Assumptions
pi0 = 93914/7166167
alpha = 1/100
alpha0 = 0
m = 300
beta = seq(from = 0, to = 30, length.out = m)/100
res_moment = res_smle = matrix(NA, m, 3)
V = mean(covid19_austria$weights^2)

for (i in 1:m){
  # Moment estimator
  inter = moment_estimator(R3 = R3w, n = n, pi0 = pi0,
                           alpha = alpha, alpha0 = alpha0,
                           beta = beta[i], V = V)

  res_moment[i,] = c(inter$estimate, inter$ci_asym)

  # Survey MLE
  inter = survey_mle(R = R1w + R3w, n = n, pi0 = pi0,
                     alpha = alpha, alpha0 = alpha0,
                     beta = beta[i], V = V)

  res_smle[i,] = c(inter$estimate, inter$ci_asym)
}

cols = c("#F8766DFF", "#00BFC4FF")
cols2 = c("#F8766D1F", "#00BFC41F")

plot(NA, xlim = 100*range(beta), ylim = c(1, 4.25), axes = FALSE, ann = FALSE)
grid()
box()
axis(1)
axis(2)
mtext(expression(paste(beta, " (%)")), side = 1, line = 3, cex = 1.15)
mtext("Prevalence (%)", side = 2, line = 3, cex = 1.15)
abline(h = 100*pi0, lwd = 2, lty = 2)
abline(h = 100*pi0, lwd = 2, lty = 2)

text(2.5, 1.18, expression(pi[0]), cex = 1.15)

legend("topleft", c("MME", "95% CI",
                    "Survey MLE", "95% CI"),
       bty = "n", col = c(cols[1], cols2[1],cols[2], cols2[2]),
       lwd = c(3, NA, 3, NA), pch = c(NA, 15, NA, 15),
       pt.cex = 2.5)
lines(100*beta, 100*res_moment[,1], lwd = 3, col = cols[1])
polygon(100*c(beta, rev(beta)),
        100*c(res_moment[,2], rev(res_moment[,3])),
        col = cols2[1], border = NA)

lines(100*beta, 100*res_smle[,1], lwd = 3, col = cols[2])
polygon(100*c(beta, rev(beta)),
        100*c(res_smle[,2], rev(res_smle[,3])),
        col = cols2[2], border = NA)
```

# Simulation Study

The results summarized in Section 5 @guerrier2020prevalence can be replicated as follows:

1. Run the file `pempi/simulations/simulation_script.R` which should take a couple of hours on a standard laptop (with $5 \times 10^4$ Monte Carlo replications). This allows to generate the file `pempi/simulations/simulations.RData`.
2. Run the file `pempi/simulations/figures.Rnw` (which reads `pempi/simulations/simulations.RData`) to generate the figures associated to our simulation study.

# References







